Effects of Water Permeability on the Hydrodynamics of a Vesicle Under Flow and Confinement

### 2020-12-17 vesicle_code/docs/permeableVes/main.tex
fixed merge conflicts

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Quiescent Flow} 
\begin{figure*}[htp]
  \centering
  \ifTikz
  \input{figures/relaxationComposite.tikz}
  \fi
  \caption{\label{fig:relaxationComposite} (a): Snapshots of four
  semi-permeable vesicles with $\beta = 10^{-3}$ submerged in a
  quiescent flow. The fluid flux along the red regions is into the
  vesicle, and the fluid flux along the blue regions is out of the
  vesicle. In all cases, the vesicle reaches a steady state circular
  shape. (b): The reduced area of each of the simulations. The
  particular snapshots in (a) occur at the marks along the curve.}
\end{figure*}

Our analysis shows that a semi-permeable vesicle in a quiescent flow
relaxes to a circle at a rate that is proportional to $\beta$, and the
steady state tension is $\Lambda = 2\pi^2/L^2$, see Supplementary Material. 
Our analysis further shows that the time for 
a semipermeable vesicle to deviate from its initial reduced area and to reach its
steady state is proportional to $\beta^{-1}$. For this reason, our numerical results below 
are reported in terms of the scaled time $\beta t$.

@@

For the vesicle with an initial reduced area ($\alpha(0)=0.12$ in the bottom row) we observe that vesicle deflates (blue regions) first due to the high interior capillary pressure.  Over time
all vesicles inflate and as expected, each vesicle eventually
reaches a circular shape with reduced area $\alpha=1$. 
%These regions are denoted in blue and are only present at early times. 
Finally, the vesicles's reduced area do not 
change by much at early times, and intermediate shapes of semi-permeable vesicles
resemble the shape of an impermeable vesicle. This can be explained by
considering the two terms that contribute to the vesicle velocity: (i)
the permeability velocity, $\beta(\ff \cdot \nn) \nn$, and (ii) the
force balance velocity, $\SS[\ff]$. Since $\beta = 10^{-3}$, the
permeability velocity is much smaller than the force balance velocity at
early times, and the semi-permeable vesicle dynamics resembles those of
an impermeable vesicle. As the force balance velocity decreases, the
permeability velocity dominates, and the vesicle begins to inflate to a
circle with no vanishing force and velocity. {\bf YNY: We should put a tick on the time when this transition occurs for each of the vesicles}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Planar Shear Flow}
\begin{figure}[htp]
  \centering
  \ifTikz
  \input{figures/shearComposite.tikz}
  \fi
  \caption{\label{fig:shearComposite} (a) The reduced area of a
  semi-permeable vesicle with $\beta = 10^{-3}$ initialized with four
  different reduced areas in a shear flow with two different flow rates.
  (b) The steady state shape of a semi-permeable vesicle with varying
  flow rates and semi-permeability rates. The circled vesicles
  correspond to the simulations in part (a). {\bf YNY: should this be two-column wide?}}
\end{figure}

@@

For a variety of semi-permeability rates
and flow rates, a semipermeable vesicle reaches a steady equilibrium as in Figure~\ref{fig:shearComposite}(a).
Since the final shape is independent of the initial
reduce area,  we initialize the vesicle with reduced
area $\nu = 0.65$ for all simulations in Figure~\ref{fig:shearComposite}(b), where we summarize the steady state vesicle shape and its reduced
area. We see that the flow
rate significantly affects the final vesicle shape, but the
semi-permeability rate has a minor effect. Instead, as we saw in the
quiescent analysis, $\beta$ sets a time scale since the time required
for a vesicle to reach its steady state configuration is proportional to
$\beta^{-1}$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Poiseuille Flow}
\begin{figure*}[htp]
  \ifTikz
  \centering
  \input{figures/parabolicComposite.tikz}
  \fi
  \caption{\label{fig:parabolicComposite} (a) The steady state shape of
  a semi-permeable vesicle submerged in a Poiseuille flow with varying
  initial reduced areas and flow rates. (b) The steady state inclination
  angle for four different flow velocities. The steady state shapes are
  superimposed. (c) The reduced area of the slipper formed with the flow
  rate of 800 $\mu m/s$. (d) The vesicle shape shown with the background
  imposed flow. The channel width is 12.5 times larger than the vesicle
  radius. The corresponding reduced area of each vesicle is indicated by
  the marks in plot (c).}
\end{figure*}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Semi-Permeable Vesicle in Stenosis}
Cells experience stretching and release messengers when they pass through capillary vessels \cite{Wan2008_PNAS,ForsythWan2011_PNAS,Pak2015_PNAS}.
Many cells exhibit extraordinary flexibility as they go through narrow channels \cite{AuStoreyMoore2016_PNAS}, and adjustment of area to volume ratio may be a key factor for successful passages.
Here we quantify the effect of water permeability on a vesicle going through a stenosis (a long microfluidic channel with a cross section smaller than the cell radius).
Based on the findings in 
%We consider the effect of stenosis on a semi-permeable vesicle. Inspired
%by an experiment that measures the effects of cells on the dynamical
%pressure-drop variations along a micrometer-sized
\cite{abk-fai-sto2006}, we first examining the effect of
semi-permeability on the pressure fluctuation across a stenosed
geometry. In this experiment, a pressure drop is used to squeeze a
red blood cell (RBC) through a constriction with diameter $5\mu$m and
length $45\mu$m. We set up a similar computational geometry with a
Poiseuille flow imposed at the inlet and outlet
(Figure~\ref{fig:stenosisComposite}(a)). The flow rate is chosen so that
the residency time of the RBC in the constriction agrees with the
experimental results. When the RBC enters the constriction, the result
is an increase in the pressure drop required to maintain a constant flow
rate, and the difference between these pressures is called the excess
pressure. We consider both an impermeable and semi-permeable vesicle
with initial reduced area $\nu = 0.65$ passing through a constricted
geometry and compute the pressure drop between the inlet and outlet of
the constriction. In Figure~\ref{fig:stenosisComposite}(c), we plot the
excess pressures. The impermeable case agrees with experimental
results~\cite{abk-fai-sto2006} (see Figure 3), and the results indicate
that less pressure is required to maintain a constant flow rate when the
vesicle is semi-permeable. Our simulations indicate the
semi-permeability does affect the excess pressure, and the vesicle's
reduced area reduces by about 6.1\%
(Figure~\ref{fig:stenosisComposite}(d)). Since the area does decrease at
a constant rate we expect that a larger amounts of deflation is possible
if the vesicle is stenosed for a longer period of time. Therefore, in
Figure~\ref{fig:stenosisComposite}(b), we consider a vesicle passing
through a much longer constriction, and we observe that the reduced area
decreases by 13.2\% (Figure~\ref{fig:stenosisComposite}(e)).

\begin{figure*}[htp]
  \ifTikz
  \input{figures/stenosisComposite.tikz}
  \fi
  \caption{\label{fig:stenosisComposite} (a) A semi-permeable vesicle
  passing through a constricted geometry similar to the experimental
  device in~\cite{abk-fai-sto2006}. (b) A semi-permeable vesicle passing
  through a constricted geometry that is over five times longer. (c) The
  excess pressure of an impermeable and semi-permeable vesicle in the
  short geometry in (a). (d) The reduced area of the semi-permeable
  vesicle in the short geometry in (a). (e) The reduced area of the
  semi-permeable vesicle in the long geometry in (b). In parts (c), (d),
  and (e), the horizontal axis is the x-coordinate of the vesicle's
  center of mass.}
\end{figure*}

%
%\begin{figure}[htp]
%%  \ifTikz
%%  \input{figures/stenosisGeom.tikz}
%%  \fi
%  \includegraphics[width=\linewidth]{figures/stenosisGeom.pdf}
%  \caption{\label{fig:stenosisGeom} A single semi-permeable vesicle
%  passing through a constriction. The geometry and parameters are chosen
%  to replicate the experiments of Abkarian et
%  al.~\cite{abk-fai-sto2006}.}
%\end{figure}
%
%\begin{figure}[htp]
%%  \ifTikz
%%  \input{figures/stenosisPressureFluctuation.tikz}
%%  \fi
%  \includegraphics[width=\linewidth]{figures/stenosisPressureFluctuation.pdf}
%  \caption{\label{fig:stenosisPressure} The excess pressure across the
%  constriction in Figure~\ref{fig:stenosisGeom}. The black dashed lines
%  denote the start and end of the constriction. The results are
%  qualitative agreement with~\cite{abk-fai-sto2006}.}
%\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Semi-Permeable Vesicle in a Contracting Geometry}

\begin{figure*}[htp]
  \centering
%  \ifTikz
  \input{figures/contractingComposite.tikz}
%  \fi
  \caption{\label{fig:contractingComposite} (a) A semi-permeable vesicle
  passing through a contracting geometry similar to the experimental
  device in~\cite{}. The narrowest part of the channel is $1.6\mu$m, and
  the vesicle height is initially $10\mu$m.}
\end{figure*}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Quiescent Flow} 
\begin{figure*}[htp]
  \centering
  \ifTikz
  \input{figures/relaxationComposite.tikz}
  \else
  \includegraphics{figures/relaxationComposite.pdf}
  \fi
  \caption{\label{fig:relaxationComposite} (a): Snapshots of four
  semi-permeable vesicles with $\beta = 10^{-3}$ submerged in a
  quiescent flow. The fluid flux along the red regions is into the
  vesicle, and the fluid flux along the blue regions is out of the
  vesicle. In all cases, the vesicle reaches a steady state circular
  shape. (b): The reduced area of each of the simulations. The
  particular snapshots in (a) occur at the marks along the curve.}
\end{figure*}

Our analysis shows that a semi-permeable vesicle in a quiescent flow
relaxes to a circle at a rate that is proportional to $\beta$, and the
steady state tension is $\Lambda = 2\pi^2/L^2$ (see Supplementary
Material). Our analysis further shows that the time for a semipermeable
vesicle to deviate from its initial reduced area and to reach its steady
state is proportional to $\beta^{-1}$. For this reason, our numerical
results below are reported in terms of the scaled time $\beta t$.

@@

For the vesicle with an initial reduced area ($\alpha(0)=0.12$ in the bottom row) we observe that vesicle deflates (blue regions) first due to the high interior capillary pressure.  Over time
all vesicles inflate and as expected, each vesicle eventually
reaches a circular shape with reduced area $\alpha=1$. 
%These regions are denoted in blue and are only present at early times. 
Finally, the vesicles's reduced area do not steady state tension is
$\Lambda = 2\pi^2/L^2$. Therefore, the times for
the vesicle to deviate from its initial reduced area and to reach its
steady state are proportional to $\beta^{-1}$.

In Figure~\ref{fig:relaxationComposite}, we examine the transient
dynamics of four vesicles with $\beta = 10^{-3}$ and a variety of
initial configurations. As expected, each vesicle eventually
reaches a circular shape with reduced area $\alpha=1$. Next, we observe
that even though the vesicle is inflating to a circle, meaning that
there is a net flux into the vesicle, there are regions where the flux
is out of the vesicle. These regions are denoted in blue and are only
present at early times. Finally, the vesicles's reduced area do not
change by much at early times, and intermediate shapes of semi-permeable
vesicles resemble the shape of an impermeable vesicle. This can be
explained by considering the two terms that contribute to the vesicle
velocity: (i) the permeability velocity, $\beta(\ff \cdot \nn) \nn$, and
(ii) the force balance velocity, $\SS[\ff]$. Since $\beta = 10^{-3}$,
the permeability velocity is much smaller than the force balance
velocity at early times, and the semi-permeable vesicle dynamics
resembles those of an impermeable vesicle. As the force balance velocity
decreases, the permeability velocity dominates, and the vesicle begins
to inflate to a circle with no vanishing force and velocity. {\bf YNY:
We should put a tick on the time when this transition occurs for each of
the vesicles}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Planar Shear Flow}
\begin{figure}[htp]
  \centering
  \ifTikz
  \input{figures/shearComposite.tikz}
  \else
  \includegraphics{figures/shearComposite.pdf}
  \fi
  \caption{\label{fig:shearComposite} (a) The reduced area of a
  semi-permeable vesicle with $\beta = 10^{-3}$ initialized with four
  different reduced areas in a shear flow with two different flow rates.
  (b) The steady state shape of a semi-permeable vesicle with varying
  flow rates and semi-permeability rates. The circled vesicles
  correspond to the simulations in part (a). {\bf YNY: should this be two-column wide?}}
\end{figure}

@@

For a variety of semi-permeability rates and flow rates, a semipermeable
vesicle reaches a steady equilibrium as in
Figure~\ref{fig:shearComposite}(a).  Since the final shape is
independent of the initial reduced area,  we initialize the vesicle with
reduced area $\alpha = 0.65$ for all simulations in
Figure~\ref{fig:shearComposite}(b), where we summarize the steady state
vesicle shape and its reduced area. We see that the flow rate
significantly affects the final vesicle shape, but the semi-permeability
rate has a minor effect. Instead, as we saw in the quiescent analysis,
$\beta$ sets a time scale since the time required for a vesicle to reach
its steady state configuration is proportional to $\beta^{-1}$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Poiseuille Flow}
\begin{figure*}[htp]
  \ifTikz
  \centering
  \input{figures/parabolicComposite.tikz}
  \else
  \includegraphics{figures/parabolicComposite.pdf}
  \fi
  \caption{\label{fig:parabolicComposite} (a) The steady state shape of
  a semi-permeable vesicle submerged in a Poiseuille flow with varying
  initial reduced areas and flow rates. The fluid flux along the red
  regions is into the vesicle, and the fluid flux along the blue regions
  is out of the vesicle.  (b) The steady state inclination angle for
  four different flow velocities. The steady state shapes are
  superimposed. (c) The reduced area of the slipper formed with the flow
  rate of 800 $\mu m/s$. (d) The vesicle shape shown with the background
  imposed flow. The channel width is 12.5 times larger than the vesicle
  radius. The corresponding reduced area of each vesicle is indicated by
  the marks in plot (c).}
\end{figure*}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Semi-Permeable Vesicle in Stenosis}
Cells experience stretching and release messengers when they pass through capillary vessels \cite{Wan2008_PNAS,ForsythWan2011_PNAS,Pak2015_PNAS}.
Many cells exhibit extraordinary flexibility as they go through narrow channels \cite{AuStoreyMoore2016_PNAS}, and adjustment of area to volume ratio may be a key factor for successful passages.
Here we quantify the effect of water permeability on a vesicle going through a stenosis (a long microfluidic channel with a cross section smaller than the cell radius).
Based on the findings in 
%We consider the effect of stenosis on a semi-permeable vesicle. Inspired
%by an experiment that measures the effects of cells on the dynamical
%pressure-drop variations along a micrometer-sized
\cite{abk-fai-sto2006}, we first examining the effect of
semi-permeability on the pressure fluctuation across a stenosed
geometry. In this experiment, a pressure drop is used to squeeze a
red blood cell (RBC) through a constriction with diameter $5\mu$m and
length $45\mu$m. We set up a similar computational geometry with a
Poiseuille flow imposed at the inlet and outlet
(Figure~\ref{fig:stenosisComposite}(a)). The flow rate is chosen so that
the residency time of the RBC in the constriction agrees with the
experimental results. When the RBC enters the constriction, the result
is an increase in the pressure drop required to maintain a constant flow
rate, and the difference between these pressures is called the excess
pressure. We consider both an impermeable and semi-permeable vesicle
with initial reduced area $\alpha = 0.65$ passing through a constricted
geometry and compute the pressure drop between the inlet and outlet of
the constriction. In Figure~\ref{fig:stenosisComposite}(c), we plot the
excess pressures. The impermeable case agrees with experimental
results~\cite{abk-fai-sto2006} (see Figure 3), and the results indicate
that less pressure is required to maintain a constant flow rate when the
vesicle is semi-permeable. Our simulations indicate the
semi-permeability does affect the excess pressure, and the vesicle's
reduced area reduces by about 6.1\%
(Figure~\ref{fig:stenosisComposite}(d)). Since the area does decrease at
a constant rate we expect that a larger amounts of deflation is possible
if the vesicle is stenosed for a longer period of time. Therefore, in
Figure~\ref{fig:stenosisComposite}(b), we consider a vesicle passing
through a much longer constriction, and we observe that the reduced area
decreases by 13.2\% (Figure~\ref{fig:stenosisComposite}(e)).

\begin{figure*}[htp]
%  \ifTikz
  \input{figures/stenosisComposite.tikz}
%  \fi
  \caption{\label{fig:stenosisComposite} (a) A semi-permeable vesicle
  passing through a constricted geometry similar to the experimental
  device in~\cite{abk-fai-sto2006}. (b) A semi-permeable vesicle passing
  through a constricted geometry that is over five times longer. (c) The
  excess pressure of an impermeable and semi-permeable vesicle in the
  short geometry in (a). (d) The reduced area of the semi-permeable
  vesicle in the short geometry in (a). (e) The reduced area of the
  semi-permeable vesicle in the long geometry in (b). In parts (c), (d),
  and (e), the horizontal axis is the x-coordinate of the vesicle's
  center of mass.}
\end{figure*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Semi-Permeable Vesicle in a Contracting Geometry} 
We expose a single semi-permeable vesicle to even larger deformations by
simulating its hydrodynamics in a contracting microfluidic geometry.
Similar to~\cite{wu2015critical}, we consider a channel whose width
gradually reduces from $16\mu$m to $1.6\mu$m and then immediately opens
again to $16\mu$m (Figure~\ref{fig:contractingComposite}). We impose a
Poiseuille flow at the inlet and outlet with a constant maximum flow
rate of $U_{\max}$. We initialize a semi-permeable vesicle with $\alpha
= 0.65$ that spans $10\mu$m of the total channel width and investigate
the effects of the permeability rate and the maximum flow velocity. We
color the vesicle according to the sign of its flux---fluid flows into
the vesicle in red regions and out of the vesicle in blue regions.

In Figures~\ref{fig:contractingComposite}a
and~\ref{fig:contractingComposite}b, we plot snapshots of vesicles with
different permeability rates, and Figure~\ref{fig:contractingComposite}c
shows that the reduced area as a function of the vesicles' location. For
sufficiently large permeability rates, the vesicle initially inflates,
but then deflates as it passes through the constriction. However, if the
semi-permeability rate is too low, the vesicle is too large when it
reaches the narrowest part of the channel, and it is unable to pass. In
Figures~\ref{fig:contractingComposite}d
and~\ref{fig:contractingComposite}e, we plot snapshots of vesicles with
different imposed maximum velocities, and
Figure~\ref{fig:contractingComposite}f shows that the reduced area as a
function of the vesicles' location. At low velocities, the vesicle
inflates similar to the unbounded parabolic flow example, and it is
unable to pass through the constrictions. However, higher flow rates
result in additional deflation and the vesicle passes through the
contraction.

To understand long term effects of a vesicle passing through a
contracting geometry, in Figures~\ref{fig:contractingComposite}g we
simulate a vesicle passing six times through the contracting geometry.
In Figure~\ref{fig:contractingComposite}h, we show that the vesicle
deflates each time it passes through the constriction followed by a
rapid inflation. The total area loss over the history of the vesicle is
almost 50\%. In Figure~\ref{fig:contractingComposite}i, we show the
vesicle velocity as a function of the vesicle's center of mass. As the
vesicle passes through the contraction, it accelerates followed by a
rapid deacceleration once it has passed through the contraction.

\begin{figure*}[htp]
  \centering
  \ifTikz
  \input{figures/contractingComposite.tikz}
  \else
  \includegraphics{figures/contractingComposite.pdf}
  \fi
  \caption{\label{fig:contractingComposite} (a) A semi-permeable vesicle
  passing through a contracting geometry similar to the experimental
  device in~\cite{}. The narrowest part of the channel is $1.6\mu$m, 
  the vesicle height is initially $10\mu$m, and the maximum background
  flow velocity is $316\mu$m/s. The fluid flux along the red regions is
  into the vesicle, and the fluid flux along the blue regions is out of
  the vesicle.}
\end{figure*}
